Random Forests", OAs

### 2024-01-08 src/oa_RF.Rmd
refined model OA

@@

@@

cities <- geo(city.names, no_query = F, method = "arcgis")

# retail centres
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
#retail <-readOGR(geo.path) 
retail <- st_read(geo.path)

#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
#   as.data.frame() %>% 
#   add_rownames(var = "id") #%>%
retail.major.cetres.help <- st_centroid(retail) %>% 
  st_transform(4326) %>%
  st_coordinates() %>% 
  as.data.frame() %>% 
  add_rownames(var = "id")

# retail.major.cetres <- retail.major.cetres.help %>% 
#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
#   st_transform(4326) %>%
#   st_coordinates() %>%
#   as_tibble() %>% 
#   bind_cols(retail.major.cetres.help$id) %>% 
#   rename(id = '...3') %>% 
#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
retail.major.cetres <- retail.major.cetres.help %>% 
  # as_tibble() %>% 
  # bind_cols(retail.major.cetres.help$id) %>% 
  # #rename(id = '...3') %>% 
  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
```

Distance to cities

@@

city.names <- city.names %>% stringr::str_remove(", UK")
names(dist) <- city.names

@@

dist <- transform(dist, dist = do.call(pmin, dist[-1]))

@@

# calculate area
oa.gb$area <- st_area(oa.gb) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
  rename(oa11cd=id) %>% 
  left_join(oa.gb, by = c("oa11cd" = "id")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,
         density = n / area) %>% 
  relocate(density, .after = area) %>% 
  dplyr::select(-geometry)

@@

Distance to retail centres

@@

# retail.names <- retail.major.cetres$RC_Name
# names(dist.retail) <- retail.names

@@

#spatial join retail centroids into OA
retail.major.cetres.help <- st_centroid(retail) %>% 
  st_transform(4326)

# lookup obj with retail id and OA id
help <- st_join(retail.major.cetres.help, oa.gb) %>% 
  as_tibble %>% 
  mutate(dist.retail.name = row_number(),
         dist.retail.name = paste0("V", dist.retail.name)) %>% 
  dplyr::select(id, dist.retail.name) %>% 
  rename(oa11cd.retail = id)

dist.retail <- dist.retail %>% 
  left_join(help, by = c('dist.retail.name' = 'dist.retail.name')) %>% 
  dplyr::select(-dist.retail.name)
sapply(dist.retail, function(x) sum(is.na(x))) 
# 1 retail centre is missing, because the centroid falls in the water

dist.retail %>% filter(is.na(oa11cd.retail))
missing.retail.centre <- retail.major.cetres.help %>% filter(RC_ID=="RC_EW_24")
st_nearest_feature(missing.retail.centre, oa.gb)
oa.gb[91986,] # missing OA: E00094765

dist.retail <- dist.retail %>%
  mutate(oa11cd.retail = ifelse(is.na(oa11cd.retail), "E00094765", oa11cd.retail))

# dist.retail %>% distinct(dist.retail.name) #6405
# help %>% distinct(dist.retail.name) # 6423 
# dist.retail %>% distinct(oa11cd.retail) #6217 as there are multiple centres in an OA

# Join with complete panel
df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))

sapply(df, function(x) sum(is.na(x)))
df %>% filter(is.na(oa11cd.retail)) %>% distinct(oa11cd) #493 rows or 29 OA don't have oa11cd.retail

# df <- df %>% dplyr::select(-dist.retail.x, -oa11cd.retail.x) %>% 
#   rename(dist.retail = dist.retail.y,
#          oa11cd.retail = oa11cd.retail.y)
```

## df for RF

```{r}
## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
#df <- df %>% filter(!is.na(area))

# London = Islington 
cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
cities.sf <- st_set_crs(cities.sf, 4326)

# help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
#   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)

sf_use_s2(FALSE)
help1 <- st_filter( oa.gb, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Edinburgh",
                    "Glasgow"))

# too slow
#dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]

# This efficiently returns the name of the column with the shortest distance
dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])

help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
  rename(nearest.city.oa11cd = id) %>% 
  dplyr::select(oa11cd, nearest.city.oa11cd)

help.retail <- dist.retail

#####save this#####
sf_use_s2(FALSE)
oa.c <- st_centroid(oa.gb)
kn <- knearneigh(oa.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)
##################

# lags
df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(oa11cd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("oa11cd" = "oa11cd"))

# n nearest city
df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
                         rename(n.nearest.city = n), 
                       by = c("nearest.city.oa11cd" = "oa11cd",
                              "year" = "year")) 
# n London 
df <- df %>% left_join(df %>% 
                         ungroup() %>% 
                         filter(oa11cd == "E00176659") %>% 
                         dplyr::select(year, n) %>% 
                         rename(n.London = n)) %>% 
  group_by(oa11cd) %>% 
  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
  ungroup()

# n nearest retail
df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
                         rename(n.nearest.retail = n), 
                       by = c("oa11cd.retail" = "oa11cd",
                              "year" = "year")) %>% 
  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))

# Lookup: OA to regions
# NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)

df <- df %>% left_join(lookup.region, by = c("oa11cd" = "OA11CD")) %>% 
  relocate(c(RGN11NM, RGN11CD), .after=year) %>% 
  mutate(help = substring(oa11cd, 1, 1),
         RGN11CD = ifelse(help=="E", RGN11CD,
                         ifelse(help=="W", "Wales",
                                ifelse(help == "S", "Scotland", NA))),
         RGN11NM = ifelse(help=="E", RGN11NM,
                         ifelse(help=="W", "Wales",
                                ifelse(help == "S", "Scotland", NA)))) %>% 
  dplyr::select(-help)

# North/South
#lookup.region %>% distinct(RGN21NM, .keep_all = T)
south <- c("London", "South West", "South West", "East of England")
df <- df %>% mutate(south = ifelse(RGN11NM %in% south, 1, 0))

# For growth:
df <- df %>% mutate(growth = log(n/n.lag),
                    abs.growth = n - n.lag)

# regional identifier
df <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
  mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
  ungroup() %>% 
  dplyr::select(-help)

sapply(df, function(x) sum(is.na(x)))

# remove not used columns
df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
                        "Leicester", "nearest.city.oa11cd"))

# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df <- read_csv(out.path)
# df %>% write_csv(file = out.path)

@@

```{r}
london.help <- df %>% group_by(RGN11NM, year) %>% 
  summarise(n.London = sum(n)) %>% 
  filter(RGN11NM == "London") %>%
  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
  ungroup() %>% 
  dplyr::select(-RGN11NM)

#n.nearest.retail IS WRONG. FIX UPSTREAM
df <- df %>% 
  rename(oa11cd.retail = n.nearest.retail) %>% 
  dplyr::select(-n.nearest.retail.lag) %>% 
  left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
                         rename(n.nearest.retail = n), 
                       by = c("oa11cd.retail" = "oa11cd",
                              "year" = "year")) %>% 
  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))

df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
  left_join(london.help, by = c('year' = 'year'))

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996) %>% 
  dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #growth, abs.growth, RGN11NM,

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # 9 after adding South. It should be RHS variables + 2

@@

start.time <- Sys.time()
model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                     n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

@@

print(model.all)
# Random Forest 
# 
# 3716736 samples
#      10 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 2926910, 2926924, 3136005, 2926924, 3136005, 2926924, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE      Rsquared   MAE      
#    2    variance    3.865809  0.5809699  0.5313035
#    2    extratrees  3.730542  0.6168954  0.5884731
#    6    variance    3.761356  0.6063340  0.5366004
#    6    extratrees  3.612670  0.6273041  0.5331196
#   10    variance    3.929176  0.5775066  0.5449132
#   10    extratrees  3.617598  0.6278302  0.5403551
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 6, splitrule =
#  extratrees and min.node.size = 5.

@@

path.out <- paste0(path, "/outputs/rf/varimp_OA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)

path.out <- paste0(path, "/data/temp/oa_rf/all_OAs/all_OAs.RData")
save(model.all, file = path.out)

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South

@@

  model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df %>% filter(RGN11CD==i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  #assign(paste0("model.all",as.character(i)), model.all)

  file <- paste0(path, "/data/temp/oa_rf/regions/region_", as.character(i), ".RData")
  save(model.all, file = file)
  rm(model.all)
}

>>>>

### Cities as points

@@

I use the centroids from arcgis API as they better reflect the city centres than 
than the geometric centres of the boundaries, which I use for `sum(n)`

@@

cities <- geo(city.names, no_query = F, method = "arcgis") %>% 
  mutate(address = stringr::str_remove(address, ", UK"))

# # Replace London with Inner London centroid
# # Source: https://data.london.gov.uk/dataset/sub-regions-london-plan-consultation-2009
# london.path <- paste0(path, "/data/raw/London_shape")
# london.sf <- st_read(dsn = london.path, layer = "lp-consultation-oct-2009-subregions")
# # Spatial transformations
# london.sf <- st_transform(london.sf, 4326)  # EPSG code for WGS84
# 
# london.c <- london.sf %>% 
#   filter(Name == "Central") %>% 
#   st_centroid() %>% 
#   st_coordinates() %>%
#   as.data.frame() %>% 
#   rename(lat = Y, long = X) %>% 
#   mutate(address = "London, UK") %>% 
#   dplyr::select(address, lat, long)
# 
# cities <- cities %>% bind_rows(london.c) %>% 
#   slice(2:n()) %>% # to remove the old London centroid based on the cities geometry
#   mutate(address = stringr::str_remove(address, ", UK"))
```

### City boundaries

# cities
city.names <- c(
  "London", 
  "Birmingham",
  "Glasgow",
  "Liverpool",
  "Bristol",
  "Manchester",
  "Sheffield",
  "Leeds",
  "Edinburgh",
  "Leicester"
  )

# City boundaries
# Source: https://geoportal.statistics.gov.uk/search?collection=Dataset&sort=name&tags=all(BDY_TCITY%2CDEC_2015)
path.in <- paste0(path, "/data/raw/Major_Towns_and_Cities_Dec_2015_Boundaries_V2_2022_121402779136482658.geojson")

city.names <- city.names %>% stringr::str_remove(", UK")

city.boundaries.ew <- st_read(path.in) %>% 
  st_transform(4326) %>% 
  filter(TCITY15NM %in% city.names) %>% 
  dplyr::select(OBJECTID, TCITY15CD, TCITY15NM) %>% 
  rename(code = TCITY15CD,
         name = TCITY15NM)

# Source: https://geoportal.statistics.gov.uk/
path.in <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
city.boundaries.sc <- st_read(path.geo) %>% 
  st_transform(4326) %>% 
  filter(grepl('Edinburgh|Glasgow', LAD21NM)) %>% 
  dplyr::select(OBJECTID, LAD21CD, LAD21NM) %>% 
  rename(code = LAD21CD,
         name = LAD21NM)

# merge
city.boundaries <- rbind(city.boundaries.ew, city.boundaries.sc)

### Retail centres

```{r}
# retail centres <- 
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
#retail <-readOGR(geo.path) 
retail <- st_read(geo.path) %>% st_transform(4326)  # EPSG code for WGS84

#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
#   as.data.frame() %>% 
#   add_rownames(var = "id") #%>%
retail.major.cetres.help <- st_centroid(retail) %>% 
  #st_transform(4326) %>%
  st_coordinates() %>% 
  as.data.frame() %>% 
  add_rownames(var = "id")

# retail.major.cetres <- retail.major.cetres.help %>% 
#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
#   st_transform(4326) %>%
#   st_coordinates() %>%
#   as_tibble() %>% 
#   bind_cols(retail.major.cetres.help$id) %>% 
#   rename(id = '...3') %>% 
#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
retail.major.cetres <- retail.major.cetres.help %>% 
  # as_tibble() %>% 
  # bind_cols(retail.major.cetres.help$id) %>% 
  # #rename(id = '...3') %>% 
  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')

### Distance to cities

@@

names(dist) <- cities$address #city.names

@@

# min distance
dist <- transform(dist, dist = do.call(pmin, dist[-1]))

# City name for minimum distance 
dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])

@@

# calculate area
#oa.gb$area <- st_area(oa.gb) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
  rename(oa11cd=id)  
  # left_join(oa.gb, by = c("oa11cd" = "id")) %>% 
  # relocate(area, .after = n) %>% 
  # mutate(area = area / 1000000,
  #        density = n / area) %>% 
  # relocate(density, .after = area) %>% 
  # dplyr::select(-geometry)

@@

### Distance to retail centres

@@

#retail.names <- retail.major.cetres$RC_ID
names(dist.retail) <- retail.major.cetres$RC_ID #city.names

@@

### TO DELETE

# # spatial join retail centroids into OA
# retail.major.cetres.help <- st_centroid(retail) %>% 
#   st_transform(4326)
# 
# # lookup obj with retail id and OA id
# help <- st_join(retail.major.cetres.help, oa.gb) %>% 
#   as_tibble %>% 
#   mutate(dist.retail.name = row_number(),
#          dist.retail.name = paste0("V", dist.retail.name)) %>% 
#   dplyr::select(id, dist.retail.name) %>% 
#   rename(oa11cd.retail = id)
# 
# dist.retail <- dist.retail %>% 
#   left_join(help, by = c('dist.retail.name' = 'dist.retail.name')) %>% 
#   dplyr::select(-dist.retail.name)
# sapply(dist.retail, function(x) sum(is.na(x))) 
# # 1 retail centre is missing, because the centroid falls in the water
# 
# dist.retail %>% filter(is.na(oa11cd.retail))
# missing.retail.centre <- retail.major.cetres.help %>% filter(RC_ID=="RC_EW_24")
# st_nearest_feature(missing.retail.centre, oa.gb)
# oa.gb[91986,] # missing OA: E00094765.
# 
# dist.retail <- dist.retail %>%
#   mutate(oa11cd.retail = ifelse(is.na(oa11cd.retail), "E00094765", oa11cd.retail))
# 
# # dist.retail %>% distinct(dist.retail.name) #6405
# # help %>% distinct(dist.retail.name) # 6423 
# # dist.retail %>% distinct(oa11cd.retail) #6217 as there are multiple centres in an OA

# Join with complete panel
df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd")) 
#sapply(df, function(x) sum(is.na(x)))
```

## n for London, nearest city and retail

### TODELETE

```{r, eval=FALSE}
# ## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
# #df <- df %>% filter(!is.na(area)) 
# 
# cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
# cities.sf <- st_set_crs(cities.sf, 4326)
# 
# # help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
# #   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
# 
# sf_use_s2(FALSE)
# help1 <- st_filter( oa.gb, cities.sf) %>% 
#   #st_coordinates() %>% 
#   as_tibble() %>% 
#   dplyr::select(id) %>% 
#   mutate(cities = c("London",     # The city names have been manually checked and assigned
#                     "Leicester",
#                     "Leeds",
#                     "Sheffield",
#                     "Bristol",
#                     "Birmingham",
#                     "Manchester",
#                     "Liverpool",
#                     "Edinburgh",
#                     "Glasgow"))
# 
# # too slow
# #dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
# 
# # This efficiently returns the name of the column with the shortest distance
# dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
# 
# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
#   #rename(nearest.city.oa11cd = id) %>% 
#   dplyr::select(oa11cd, dist.city.name)
# 
# df <- df %>%  left_join(help2, by = c("oa11cd" = "oa11cd")) 
```

### n for cities

```{R}
#help.retail <- dist.retail

# n nearest city
# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
#                          rename(n.nearest.city = n), 
#                        by = c("nearest.city.oa11cd" = "oa11cd",
#                               "year" = "year"))

sf_use_s2(FALSE)
city.boundaries.oa <- st_join(oa.gb, city.boundaries, join = st_intersects) %>% 
  as_tibble() %>% 
  filter(!is.na(name)) %>% 
  dplyr::select(id, name) %>% 
  mutate(name = ifelse(name == "City of Edinburgh", "Edinburgh", name)) %>% 
  mutate(name = ifelse(name == "Glasgow City", "Glasgow", name))

city.boundaries.help <- df %>% inner_join(city.boundaries.oa, by = c("oa11cd" = "id")) %>% 
  group_by(name, year) %>% 
  summarise(n.nearest.city = sum(n)) %>% 
  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
  ungroup()

london.boundaries.help <- city.boundaries.help %>% 
  filter(name == "London") %>% 
  rename(n.London = n.nearest.city,
         n.London.lag = n.nearest.city.lag)

df <- df %>% left_join(city.boundaries.help, by = c("dist.city.name" = "name", "year" = "year"))

df <- df %>% left_join(london.boundaries.help, by = c("year" = "year"))
#sapply(df, function(x) sum(is.na(x)))
```

### DELETE n for London

```{r, eval=FALSE}

# # OLD London definitions
# # df <- df %>% left_join(df %>% 
# #                          ungroup() %>% 
# #                          filter(oa11cd == "E00176659") %>% 
# #                          dplyr::select(year, n) %>% 
# #                          rename(n.London = n)) %>% 
# #   group_by(oa11cd) %>% 
# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
# #   mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
# #   ungroup() 
# # 
# # london.help <- df %>% group_by(RGN11NM, year) %>% 
# #   summarise(n.London = sum(n)) %>% 
# #   filter(RGN11NM == "London") %>%
# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
# #   ungroup() %>% 
# #   dplyr::select(-RGN11NM) 
# 
# london.oa <- st_filter(oa.gb, london.sf %>% 
#                          filter(Name == "Central")) %>% 
#   as_tibble() %>% 
#   dplyr::select(id)
# 
# london.help <- df %>% inner_join(london.oa, by = c("oa11cd" = "id")) %>% 
#   group_by(year) %>% 
#   summarise(n.London = sum(n)) %>% 
#   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
#   ungroup()
# 
# df <- df %>% left_join(london.help, by = c("year" = "year"))
```

### n for nearest retail

```{r}
# Old definition for retail centre
# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
#                          rename(n.nearest.retail = n), 
#                        by = c("oa11cd.retail" = "oa11cd",
#                               "year" = "year")) %>% 
#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
# 
# 
# #n.nearest.retail IS WRONG. FIX UPSTREAM
# df <- df %>% 
#   rename(oa11cd.retail = n.nearest.retail) %>% 
#   dplyr::select(-n.nearest.retail.lag) %>% 
#   left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
#                          rename(n.nearest.retail = n), 
#                        by = c("oa11cd.retail" = "oa11cd",
#                               "year" = "year")) %>% 
#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
# 
# 
# df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
#   left_join(london.help, by = c('year' = 'year'))

retail.oa <- st_join(oa.gb, retail, join = st_intersects) %>% 
  as_tibble() %>% 
  filter(!is.na(RC_ID)) %>% 
  dplyr::select(id, RC_ID)

# st_join(oa.gb, retail, join = st_is_within_distance, dist = 500) %>% 
#   as_tibble() %>% 
#   filter(!is.na(RC_ID)) %>% 
#   dplyr::select(id, RC_ID)
# 
# st_filter(oa.gb, retail,  .pred = st_is_within_distance, dist = 100) %>% 
#   as_tibble() %>% 
#   filter(!is.na(RC_ID)) %>% 
#   dplyr::select(id, RC_ID)

# h <- oa.gb %>% st_intersects(retail)

# retail.oa_ %>% filter(id == "E00004828"|id == "E00004829"|id == "E00004827")
# # RC_EW_5285
# retail.oa <- retail.oa %>% add_row(id = "E00004828", RC_ID = "RC_EW_5285")
# retail.oa <- retail.oa %>% arrange(id)

retail.help <- df %>% inner_join(retail.oa, by = c("oa11cd" = "id")) %>% 
  group_by(RC_ID, year) %>% 
  summarise(n.nearest.retail = sum(n)) %>% 
  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail)) %>% 
  ungroup()

# retail.help2 <- retail.oa %>% left_join(retail.help, by = c("RC_ID" = "RC_ID")) %>% 
#   dplyr::select(-RC_ID)

df <- df %>% left_join(retail.help, by = c("dist.retail.name" = "RC_ID", "year" = "year"))
#sapply(df_, function(x) sum(is.na(x)))
```

## TODO

```{r}
# Lookup: OA to regions
# **FIXED< DELETE IN TEST** NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)

df <- df %>% left_join(lookup.region, by = c("oa11cd" = "OA11CD")) %>% 
  relocate(c(RGN11NM, RGN11CD), .after=year) %>% 
  mutate(help = substring(oa11cd, 1, 1),
         RGN11CD = ifelse(help=="E", RGN11CD,
                         ifelse(help=="W", "Wales",
                                ifelse(help == "S", "Scotland", "NI"))),
         RGN11NM = ifelse(help=="E", RGN11NM,
                         ifelse(help=="W", "Wales",
                                ifelse(help == "S", "Scotland", "NI")))) %>% 
  dplyr::select(-help)

# # North/South
# #lookup.region %>% distinct(RGN21NM, .keep_all = T)
# south <- c("London", "South West", "South West", "East of England")
# df <- df %>% mutate(south = ifelse(RGN11NM %in% south, 1, 0))

# For growth:
df <- df %>% group_by(oa11cd) %>% 
  mutate(n.lag = dplyr::lag(n),
         growth = log(n/n.lag),
                    abs.growth = n - n.lag) %>% 
  relocate(n.lag, .after = n)

# # regional identifier
# df_ <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
#   mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
#   ungroup() %>% 
#   dplyr::select(-help)

#sapply(df, function(x) sum(is.na(x)))
```

## Spatial and spatio-temporal lags

```{r}
# sf_use_s2(FALSE)
# oa.c <- st_centroid(oa.gb)
# kn <- knearneigh(oa.c, k = 5)
# knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
# knn.l <- nb2listw(knn)

path.out <- paste0(path, "/data/temp/oa_rf/all_OAs/knnl.RData")
# save(knn.l, file = path.out)
load(path.out)

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(oa11cd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year))        # duplicate line
```

@@

```{r}
# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df <- read_csv(out.path)
# df %>% write_csv(file = out.path)

# remove not used columns
df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
                        "Leicester", "dist.city.name", "dist.retail.name", "name"))

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996) %>% 
  dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, 
                n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) %>% #growth, abs.growth, RGN11NM, 
  mutate(London.v = n.London.lag / London,
         city.v = n.nearest.city.lag / dist,
         retail.v = n.nearest.retail.lag / (dist.retail + .1)) %>% 
  dplyr::select(-n.London.lag, -London, -n.nearest.city.lag, -dist, -n.nearest.retail.lag, -dist.retail)

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # It should be RHS variables + 2

@@

start.time <- Sys.time()
model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                     n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
                     #London.v + city.v + retail.v + n.l.slag + year,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

@@

print(model.all)

# Random Forest 
# 
# 3716736 samples
#      10 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 2926910, 2926924, 3136005, 2926924, 3136005, 2926924, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE      Rsquared   MAE      
#    2    variance    3.865809  0.5809699  0.5313035
#    2    extratrees  3.730542  0.6168954  0.5884731
#    6    variance    3.761356  0.6063340  0.5366004
#    6    extratrees  3.612670  0.6273041  0.5331196
#   10    variance    3.929176  0.5775066  0.5449132
#   10    extratrees  3.617598  0.6278302  0.5403551
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 6, splitrule =
#  extratrees and min.node.size = 5.

@@

path.out <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)

path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs.RData")
save(model.all, file = path.out)

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # It should be RHS variables + 2

@@

  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
                   data = df %>% filter(RGN11CD==i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  #assign(paste0("model.all",as.character(i)), model.all)

  file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
  save(model.all, file = file)
  rm(model.all)
}
